### 2018-09-06 paper/finitecomplete-failures.tex
New version of U_F test cases introduced. These are directly implementable with FDR4

In the domain of process algebras, test cases are typically represented
by processes interacting concurrently with the SUT
process~\cite{Hennessy:1988:ATP:50497}. They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed. %%%, or if no verdict could be obtained.

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events 
$\efail$ or $\epass$ always become visible, if they are the only events
$U_F(p)/s$ is ready to engage in.

The specification of $U_F(p,s)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
If branch (\ref{eq:ufa})
is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
is the last to be entered, the process either accepts another event
$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL.

The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $(\#s = p\wedge \minhits(P/s)\neq\varnothing)$. 
This is explained further in the next paragraphs.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and
$\minhits(P/s)$ are finite. We further recall that $\minhits(P/s)$ may be
empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
is undefined. Therefore, the associated guard condition contains the requirement
$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken.
Recall that $\minhits(P/s)=\varnothing$ implies  $\minaccs(P/s) = \{\varnothing\}$,
so branches (\ref{eq:ufa}) and (\ref{eq:ufb}) can be taken in this situation.
%%%\fixme{alcc: Oops. This is undefined in CSP.}

Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
a length $p$. If, after having already run through $s\in\trc(P)$ with
$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$,
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
this is revealed by $\varnothing$ being  its only minimal acceptance.
In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
the external choice in $U_F(p,s)$).

If the length of $s$ is still less than $p$, the test accepts any event from
the initials $[P/s]^0$ and continues recursively as $U_F(p,s.e)$ in
branch~(\ref{eq:ufc}). A test of this type is called \emph{adaptive}, because
it accepts any legal behaviour of the SUT and adapts its consecutive
behaviour to the event selected by the SUT.

After having run through a trace of length $p$, and if
$\minaccs(P/s)\neq\{\varnothing\}$ so that also
$\minhits(P/s)\neq\varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set of $\minaccs(P/s)$ and
only offers the events contained in this set. If the SUT refuses to engage
into any of these events, this reveals a violation of failures refinement:
according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
only terminates with  $\epass$, if such an event is accepted by the SUT.

>>>>

% -------------------------------------------------------------------------
\subsubsection*{Test Definition and Basic Properties}
In the domain of process algebras, test cases are typically represented
by processes interacting concurrently with the SUT
process~\cite{Hennessy:1988:ATP:50497}. They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed. %%%, or if no verdict could be obtained.

For a given reference process $P$, its normalised transition graph 
$$
G(P) = ( N, \ii n, \Sigma, t : N\times\Sigma \pfun N, r : N \fun \mathbb{P}\mathbb{P}(\Sigma)),
$$ 
and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%%
%\begin{eqnarray}
%U_F(p) & = & U_F(p,\varepsilon)
%\\
%U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
%\label{eq:ufa}
%\\ & & \extchoice \nonumber
%\\ & & (\minaccs(P/s) = \{ \varnothing \})    \&   \big( \epass \then \Stop \big)
%\label{eq:ufb}
%\\ & & \extchoice \nonumber
%\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
%\label{eq:ufc}
%\\ & & \extchoice \nonumber
%\\ & & (\#s = p \wedge \minhits(P/s)\neq\varnothing) \&  \nonumber 
%\\ & & \taba \big( \Intchoice_{H\in\minhits(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
%\label{eq:ufd}
%\end{eqnarray}
%
%
\begin{eqnarray}
U_F(p) & = & U_F(p,0,\ii n)
\\
U_F(p,k,n) & = & \big(e:(\Sigma - [n]^0)  \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & (\minhits(n) =   \varnothing  )    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (k < p) \& \big(e:[n]^0   \then U_F(p,k+1,t(n,e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (k = p \wedge \minhits(n)\neq\varnothing) \&  \nonumber 
\\ & & \taba \big( \Intchoice_{H\in\minhits(n)} (e:H   \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
Before explaining the intuition behind this test definition in the paragraphs below,
we state its basic properties in the following lemma, to show the relationships between
$U_F(p)$ and the reference process $P$ from which the former has been derived.

\begin{lemma}\label{lemma:ufproperties}
If $s\in\trc(P)$ satisfies $\#s\le p$, then
$s, s.e\in\trc(U_F(p))$ for all $e\in\Sigma$, and the following properties hold.
\begin{eqnarray}
\label{eq:ifpa}
  &  & U_F(p)/s = U_F(p,\#s,G(P)/s)
\\
\label{eq:ifpb}
e\not\in [P/s]^0 & \implies & U_F(p)/s.e = (\efail\then\Stop)
\\
\label{eq:ifpc}
U_F(p)/s = U(p,\# s,n)  & \implies & [n]^0 = [P/s]^0  
\\
\label{eq:ifpd}
U_F(p)/s = U(p,\# s,n)  & \implies & \minhits(n) = \minhits(P/s)
\end{eqnarray}
\end{lemma}
\begin{proof}
We prove (\ref{eq:ifpa}) by induction over the length of $s$. For $\#s = 0$, the statement holds because $U_F(p)$ starts with the initial node $\ii n$ of $G(P)$. Suppose that 
the statement holds for all traces $s$ with length $\# s \le k < p$, so that
$U_F(p)/s = U_F(p,\#s,G(P)/s)$. Now let $s.e$ be a trace of $P$. Then $e\in [P/s]^0$. Since $[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$, we conclude that $e\in  [G(P)/s]^0$,
so $U_F(p)/s = U_F(p,\#s,G(P)/s)$ can engage into $e$ by executing branch (\ref{eq:ufc}).
 Since $t$ is the transition function of $G(P)$ and $e\in  [G(P)/s]^0$, $t(G(P)/s,e)$
 is defined, and $t(G(P)/s,e) = G(P)/s.e$ holds.
This leads to a new recursion  
$U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = U_F(p,\#s+1,G(P)/s.e)$ and proves the induction step.

To prove (\ref{eq:ifpb}), we apply (\ref{eq:ifpa}) to conclude that 
$U_F(p)/s = U_F(p,\#s,G(P)/s)$, because $s$ is a trace of $P$. Noting again that
$[G(P)/s]^0 = [P/s]^0$, this implies that $e\not\in [G(P)/s]^0$, so $U_F(p,\#s,G(P)/s)$
can engage in $e$ by entering branch (\ref{eq:ufa}). The specification of this branch
implies that 
$$
U_F(p)/s.e = U_F(p,\#s,G(P)/s)/e = (\efail\then\Stop).
$$

Statement (\ref{eq:ifpc}) follows trivially from (\ref{eq:ifpa}), because 
$[G(P)/s]^0 = [P/s]^0$ for all traces $s$ of $P$.

Finally, statement (\ref{eq:ifpd}) follows trivially from (\ref{eq:ifpa}), because,
according to (\ref{eq:minhitsGP}),
$\minhits(G(P)/s) = \minhits(P/s)$ for all traces of $P$.
\xbox
\end{proof}

Note that it is not guaranteed for $U_F(p)$ to run through the traces $s, s.e$ specified 
in Lemma~\ref{lemma:ufproperties}, if $\minhits(P/u) = \varnothing$ for some prefix $u$ of $s$: in such a case, $U_F(p)$ may stop with a $\epass$-event by entering branch (\ref{eq:ufb}). Therefore, Lemma~\ref{lemma:ufproperties} just states the existence of
$U_F(p)$-executions $s, s.e$ satisfying the properties stated there.

% --------------------------------------------------------------------------
\subsubsection*{Explanation of the Test Definition}
A test is performed by running $U_F(p)$ concurrently with any SUT process
$Q$, synchronising over $\Sigma$. So, a test execution is a trace of
the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, indicating a verdict
FAIL and PASS for the test execution, are not included in $\Sigma$. Since we
assume that $Q$ is free of livelocks, it is guaranteed that events 
$\efail$ or $\epass$ always become visible, if they are the only events
$U_F(p)/s$ is ready to engage in.

The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with   event
$\epass$. This can also be  expressed by means of a failures refinement as
defined below.
%
\begin{equation}
\label{eq:passF}
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\end{equation}
%
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\epass$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,k,n)$ executing branch (\ref{eq:ufd}),
because $(k = p\wedge \minhits(P/s)\neq\varnothing)$. 
This is explained further in the next paragraphs.

The specification of $U_F(p,k,n)$ implies that the test always stops after having engaged into a trace $s\in\trc(Q)$ of maximal length $p$ or $p+1$: 
If branch (\ref{eq:ufa})
is the last to be entered, the maximal length of $s$ is $p+1$, and the test execution
stops with $\efail$. If branch (\ref{eq:ufb}) is the last to be entered, the maximal length of $s$ is $p$, and the execution stops with $\epass$. If branch (\ref{eq:ufd})
is the last to be entered, the process either accepts another event
$e$ of some minimal hitting set $H\in\minhits(P/front(s))$. Then the final length
of $s$ is $p+1$, and the execution terminates with $\epass$. Or the test execution
$(Q\parallel[\Sigma] U_F(p))/s$ deadlocks,  the final length of $s$ is $p$,
and the execution stops without a PASS or FAIL event. Such an execution is also interpreted as FAIL, because it reveals that $(\epass\then\Stop) \not\lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma$.

We observe that the number of possible executions of $Q\parallel[\Sigma]
U_F(p)$ is finite, because the number of traces $s$ with maximal length
$(p+1)$ is finite and the sets $[n]^0$, $(\Sigma - [n]^0)$, and
$\minhits(n)$ are finite. We further recall that $\minhits(n)$ may be
empty, so that the indexed internal choice construct in (\ref{eq:ufd}) 
is undefined. Therefore, the associated guard condition contains the requirement
$\minhits(P/s)\neq\varnothing$, otherwise this branch cannot be taken,
but branches (\ref{eq:ufa}) or (\ref{eq:ufb}) can be taken in this situation.
%%%\fixme{alcc: Oops. This is undefined in CSP.}

Intuitively speaking, $U_F(p)$ is able to perform any trace $s$ of $P$, up to
a length $p$. If, after having already run through $s\in\trc(P)$ with
$\#s \le p$, an event is accepted by the SUT that is outside the initials of $P/s$
 (recall from Lemma~\ref{lemma:ufproperties} that $[n]^0 = [P/s]^0$ for $U_F(p)/s$),
the test immediately terminates with FAIL-event $\efail$. This is handled by
the branch (\ref{eq:ufa}) of the external choice in the process $U_F(p,s)$
defined above.

If $P/s$ is the $\Stop$ process or has $\Stop$ as an internal choice,
this is revealed by $\minhits(n) = \varnothing$ (recall (\ref{eq:minhitminaccempty}) and
Lemma~\ref{lemma:ufproperties}). 
In this case, the test may terminate successfully (branch (\ref{eq:ufb}) of
the external choice in $U_F(p,s)$). If $P/s$ may also nondeterministically engage into events in such a situation, branches  (\ref{eq:ufa}) and (\ref{eq:ufc}) of the test are
simultaneously enabled.

If the length of $s$ is still less than $p$, the test accepts any event $e$ from
the initials $[P/s]^0 = [G(P)/s]^0$ and continues recursively as $U_F(p,\#s+1,G(P)/s.e)$ in
branch~(\ref{eq:ufc}), this follows again from  Lemma~\ref{lemma:ufproperties}). 
A test of this type is called \emph{adaptive}, because
it accepts any legal behaviour of the SUT and adapts its consecutive
behaviour to the event selected by the SUT.

After having run through a trace of length $p$, and if
$\minhits(P/s)\neq \varnothing$, the test changes its behaviour: instead of
offering {\it all} legal events from $[P/s]^0$ to the SUT, it
nondeterministically chooses a minimal hitting set   $H\in \minhits(G(P)/s)$ 
and
only offers the events contained in $H$. If the SUT refuses to engage
into any event of $H$, this reveals a violation of failures refinement:
according to Lemma~\ref{lemma:hseta}, a conforming SUT should accept at least
one event of each minimal hitting set in $\minhits(P/s)$. Therefore, the test
only terminates with  $\epass$, if such an event is accepted by the SUT.

### 2018-09-06 paper/preliminaries.tex
New version of U_F test cases introduced. These are directly implementable with FDR4

Note that the node names including their number suffixes are generated by the
FDR tool. The numbering is generated during the normalisation procedure. So,
the node numbers do not reflect the distance from the initial node Node\_0.
\xbox
\end{example}
%
Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs
\begin{eqnarray*}
G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
\\
G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
\end{eqnarray*}
%
as established by the results  in the following lemma.
There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
by means of the maximal refusal functions of the graphs involved; and
(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
can be derived from the maximal refusal functions by means of
(\ref{eq:maxrefsminaccs}). 
%%\fixme{alcc: I think at least the report should point to proofs.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge P\ conf\ Q
  \\
  \label{eq:failrefa}
  Q\ conf\ P & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow &
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%
\begin{proof}
Recall that $P\lessdet_T Q$ is defined
as $\trc(Q)\subseteq \trc(P)$ and that $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$.
This shows (\ref{eq:trcrefa}).

\medskip
\noindent
To prove (\ref{eq:failconf}), we derive
\begin{argue}
P \lessdet_F Q 
\\
\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
\\
\Leftrightarrow \forall (s,X) \in  \fails(Q): (s,X) \in\fails(P) 
& property of $\subseteq$
\\
\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
\wedge {}
\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
\\
& Definition of failures, refusals, subset closure
\\
\Leftrightarrow \trc(Q)\subseteq\trc(P)
\wedge {}
\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
\\
& Definition of failures and traces
\\
\Leftrightarrow \trc(Q)\subseteq\trc(P)
\wedge {}
\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
\\
& Property of $\cap$ and $\subseteq$
\\
\Leftrightarrow
P\lessdet_T Q \wedge Q\ conf\ P
& Definition of $\lessdet_T$ and $conf$
\end{argue}

@@

For proving our main theorems, it is necessary to consider the \emph{product}
of normalised transition graphs. We need this only for the investigation of
corresponding traces in reference processes and processes for SUTs. So, the
labelling of nodes with maximal refusals or minimal acceptances are
disregarded in the product construction. We consider two normalised
transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
%
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
\\ & & \tabc
(n_1,e)\in\dom~t_1\wedge
(n_2,e) \in\dom~t_2    \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}
%
The following lemma is used in the proof of our main theorem.
%
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}
\fixme{Abrupt end to the section. It has more than just preliminary material.
It has some new results.}

@@

To prove ``$2 \Rightarrow 1$'', assume that 2 holds, but that $P\ conf\ Q$
does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$.
Minimising $\overline H$ yields   a minimal hitting set $H\in
\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%
We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
$\minhits(P/s) = \varnothing$.

>>>>

Note that the node names including their number suffixes are generated by the
FDR tool. The numbering is generated during the normalisation procedure. So,
the node numbers do not reflect the distance from the initial node Node\_0.
\xbox
\end{example}
%
Summarising, refinement relations between finite-state CSP processes $P, Q$ can be be
expressed by means of their normalised transition graphs
\begin{eqnarray*}
G(P) & = & ( N_P, {\ii n}_P, \Sigma, t_P : N_P\times\Sigma \pfun N_P, r_P : N_P \fun \mathbb{P}\mathbb{P}(\Sigma))
\\
G(Q) & = & ( N_Q, {\ii n}_Q, \Sigma, t_Q : N_Q\times\Sigma \pfun N_Q, r_Q : N_Q \fun \mathbb{P}\mathbb{P}(\Sigma))
\end{eqnarray*}
%
as established by the results  in the following lemma.
There, result (\ref{eq:trcrefa}) reflects trace refinement in terms of graph
traces; (\ref{eq:failconf}) expresses failures refinement in terms of traces
refinement and $conf$; (\ref{eq:failrefa}) states how $conf$ can be expressed
by means of the maximal refusal functions of the graphs involved; and
(\ref{eq:failrefb}) states the same in terms of the minimal acceptances that
can be derived from the maximal refusal functions by means of
(\ref{eq:maxrefsminaccs}). 
%%\fixme{alcc: I think at least the report should point to proofs.}
%
\begin{lemma}
  \label{lemma:tgtrcref}
  \begin{eqnarray}
  P \lessdet_T Q & \Leftrightarrow & \trc(G(Q)) \subseteq\trc(G(P))
  \label{eq:trcrefa}
  \\
  \label{eq:failconf}
  P \lessdet_F Q & \Leftrightarrow & P \lessdet_T Q \wedge Q\ conf\ P
  \\
  \label{eq:failrefa}
  Q\ conf\ P & \Leftrightarrow & \nonumber
  \forall s\in\trc(G(Q))\cap \trc(G(P)), R_Q\in r_Q(G(Q)/s):  \nonumber
  \\ & & \tabd
  \exists R_P\in r_P(G(P)/s): R_Q\subseteq R_P
  \\
  \label{eq:failrefb}
   & \Leftrightarrow &
    \forall s\in\trc(G(Q))\cap \trc(G(P)), A_Q\in\minaccs(G(Q)/s):  \nonumber
   \\ & & \tabd
  \exists A_P\in\minaccs(G(P)/s): A_P\subseteq A_Q
 \end{eqnarray}
  \xbox
\end{lemma}
%
\begin{proof}
Recall that $P\lessdet_T Q$ is defined
as $\trc(Q)\subseteq \trc(P)$ and that $\trc(P)=\trc(G(P))$ and $\trc(Q)=\trc(G(Q))$.
This shows (\ref{eq:trcrefa}).

\medskip
\noindent
To prove (\ref{eq:failconf}), we derive
\begin{argue}
P \lessdet_F Q 
\\
\Leftrightarrow \fails(Q)\subseteq \fails(Q) & definition of $\lessdet_F$
\\
\Leftrightarrow \forall (s,R) \in  \fails(Q): (s,R) \in\fails(P) 
& property of $\subseteq$
\\
\Leftrightarrow \forall (s,\varnothing) \in  \fails(Q): (s,\varnothing) \in\fails(P)
\wedge {}
\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
\\
& Definition of failures, refusals, subset closure
\\
\Leftrightarrow \trc(Q)\subseteq\trc(P)
\wedge {}
\\\tabc \forall s\in\trc(Q), R\in \refs(Q/s): R\in\refs(P/s)
\\
& Definition of failures and traces
\\
\Leftrightarrow \trc(Q)\subseteq\trc(P)
\wedge {}
\\\tabc \forall s\in\trc(Q)\cap\trc(P): \refs(Q/s) \subseteq\refs(P/s)
\\
& Property of $\cap$ and $\subseteq$
\\
\Leftrightarrow
P\lessdet_T Q \wedge Q\ conf\ P
& Definition of $\lessdet_T$ and $conf$
\end{argue}

@@

For proving our main theorems, it is necessary to consider the \emph{product}
of normalised transition graphs. We need this only for the investigation of
corresponding traces in reference processes and processes for SUTs. So, the
labelling of nodes with maximal refusals or minimal acceptances are
disregarded in the product construction. We consider two normalised
transition graphs
\[
G_i = ( N_i, \ii n_i, \Sigma, t_i : N_i\times\Sigma \pfun N_i, r_i : N_i \fun \mathbb{P}\mathbb{P}(\Sigma)),\qquad i = 1,2,
\]
over the same alphabet $\Sigma$. Their product is defined by
%
\begin{eqnarray}
G_1\times G_2 & = & (N_1\times N_2,(\ii n_1,\ii n_2), t:(N_1\times N_2)\times\Sigma\pfun (N_1\times N_2))
\\
\dom~t & = & \{ ((n_1,n_2),e)\in (N_1\times N_2)\times\Sigma~|   \nonumber
\\ & & \tabc
(n_1,e)\in\dom~t_1\wedge
(n_2,e) \in\dom~t_2    \}
\\
t((n_1,n_2),e) & = & (t_1(n_1,e),t_2(n_2,e))\ \text{for $((n_1,n_2),e)\in\dom~t$}
\end{eqnarray}
%
The following lemma is used in the proof of our main theorem.
%
\begin{lemma}\label{lemma:reachproduc}
If $G_1$ has $p$ states and $G_2$ has $q$ states, then every reachable state
$(n_1,n_2)$ of the product graph $G_1\times G_2$ can be reached by a trace
%%%$s\in\trc(G_1)\cap \trc(G_2)$  we don't need this
of maximal length $(pq-1)$.
\end{lemma}
\begin{proof}
The product graph $G_1\times G_2$ has at most $pq$ states. The empty trace $\varepsilon$
reaches its initial state $(\ii n_1,\ii n_2)$. Applying Lemma~\ref{lemma:extendV}
$(pq-1)$ times with $V=\{\varepsilon \}$ implies that $G_1\times G_2$ reaches
all of its reachable states (there are at most $pq$ of them) under
$V' = V \cup V.\Sigma\cup\dots \cup V.\Sigma^{(pq-1)}$. The maximal length of traces in
$V'$ is $(pq-1)$.
\xbox
\end{proof}

%\fixme{Abrupt end to the section. It has more than just preliminary material.
%It has some new results.}

@@

To prove ``$2 \Rightarrow 1$'', assume that 2 holds, but that $P\ conf\ Q$
does {\it not} hold. According to Lemma~\ref{lemma:tgtrcref},
(\ref{eq:failrefb}), there exists $s\in\trc(P)\cap \trc(Q)$ such that
\[
\exists A_Q\in\minaccs(G(Q)/s): \forall A_P\in\minaccs(G(P)/s): A_P\not\subseteq A_Q
\qquad (*)
\]
Let $A$ be such an acceptance set $A_Q$ fulfilling (*).
Define
\[
\overline H = \bigcup\{ A_P \setminus A~|~A_P\in\minaccs(G(P)/s) \}.
\]
Since $A_P \setminus A \neq\varnothing$ for all $A_P$ because of (*),
$\overline H$ is a hitting set of $\minaccs(G(P)/s)$ which has an  empty
intersection with $A$.
Minimising $\overline H$ yields   a minimal hitting set $H\in
\minhits(P/s)$ which is {\it not} a hitting set of $\minaccs(G(Q)/s)$, a
contradiction to Assumption~2. This completes the proof of the lemma. \xbox
\end{proof}
%
We note that $\minaccs(P) = \{ \varnothing \}$ if $P = Q\intchoice \Stop$.
Since $\Stop$ accepts nothing, its minimal acceptance is $\varnothing$, and
this carries over to $Q\intchoice \Stop$.  From (\ref{eq:failrefb}) we
conclude that $\varnothing\in\minaccs(P)$ implies $\minaccs(P) = \{
\varnothing \}$. This clarifies that $\minhits(P/s)$ is empty if, and only
if, $\minaccs(P) = \{ \varnothing \}$. The proof of Lemma~\ref{lemma:hseta}
covers the situations where $\minaccs(P/s) = \{ \varnothing \}$ and so
$\minhits(P/s) = \varnothing$. Trivially,
\begin{equation}
\label{eq:minhitminaccempty}
\minaccs(P/s) = \{ \varnothing \} \iff \minhits(P/s) = \varnothing
\end{equation}
holds.

Since minimal acceptances can be used to label the nodes of a normalised transition graph, and since $\minaccs(P/s) = \minaccs(G(P)/s)$, the notation of minimal hitting sets also
carries over to graphs: we write $\minhits(n)$ for nodes $n$ of $G(P)$ and observe that
\begin{equation}
\label{eq:minhitsGP}
\minhits(G(P)/s) = \minhits(P/s) \ \text{for all}\ s\in \trc(P).
\end{equation}
